\section{Server-Client Communication}

\subsection{Protocol}

\subsubsection{Grammar}
The following grammar will facilitate the text-based communication between the clients and the server. The server will send \texttt{StoC\_MSG} messages to the client, which will be able to send \texttt{CtoS\_MSG} messages back to the server.

\vspace{5mm}

\setlength{\parindent}{0in}

\texttt{StoC\_MSG :== (STROKE | BRD\_INFO | BRD\_DEL | USER\_INIT | BRD\_USERS) N}\\

\texttt{CtoS\_MSG :== (STROKE | SEL | BRD\_REQ | BRD\_DEL | BRD\_ALL | USER\_REQ) N}\\


\texttt{STROKE :== "stroke" S BOARD\_ID S THICK S COORDS S COLOR}\\
\texttt{COORDS :== X1 S Y1 S X2 S Y2}\\
\texttt{X1, Y1, X2, Y2 :== INT}\\
\texttt{COLOR :== BYTE S BYTE S BYTE}\\
\texttt{THICK :== ([1-9]|10)}\\ %[1-10]

\texttt{SEL :== "select" S BOARD\_ID}\\

\texttt{BRD\_REQ :== "board\_req" S NAME}\\
\texttt{BRD\_ALL :== "board\_all"}\\
\texttt{BRD\_INFO :== "board" S BOARD\_ID S NAME}\\
\texttt{BRD\_DEL :== "del" S BOARD\_ID}\\
\texttt{BRD\_CLR :== "board\_clear" S BOARD\_ID}\\
\texttt{BRD\_USERS :== "board\_users" S BOARD\_ID (S USER\_NAME)+}\\

\texttt{USER\_REQ :== "user\_req" (S USER\_NAME)?}\\
\texttt{USER\_INIT :== "you\_are" S USER\_NAME}\\

\texttt{NAME :== [$\char`\^ \backslash$r$\backslash$n]+}\\
\texttt{USER\_NAME :== [A-Za-z][A-Za-z0-9]*}\\
\texttt{BOARD\_ID :== INT}\\

\texttt{INT :== [0-9]+}\\
\texttt{BYTE :== ([0-9]{1,2}|1[0-9]{2}|2[0-4][0-9]|25[0-5])}\\ %[0-255]
\texttt{N :== "$\backslash$r?$\backslash$n"}\\
\texttt{S :== " "}\\

\setlength{\parindent}{15pt} %default

\subsubsection{Definitions and Usage}
Below is the definition for each of the messages that can be sent across the network. Note that particular messages warrant a response after processing, and those returned messages are included in the definition.

\paragraph{\texttt{STROKE}} A \texttt{STROKE} message is sent from a client to the server upon drawing a line.
\begin{enumerate}
\item A \texttt{WhiteLine} object is created for the line from (\texttt{X1},\texttt{Y1}) to (\texttt{X2},\texttt{Y2}) with thickness \texttt{THICK} (in range [1-10]) and color \texttt{COLOR} (RGB values in range [0-255]). It is then added to the \texttt{MasterBoard} corresponding to the provided \texttt{BOARD\_ID}.
\item The \texttt{STROKE} message is forwarded to all clients who are currently editing the same board, so that the line can be incorporated onto their whiteboards.
\end{enumerate}

\paragraph{\texttt{SEL}} Upon selecting a different board, the client sends a \texttt{SEL} request to the server.
\begin{enumerate}
\item The server clears all stroke messages queued to update the client's whiteboard.
\item The server disassociates the client with the current whiteboard if one is assigned and associates the client with the requested board.
\item A \texttt{BRD\_USERS} message is sent to all users of the requested board, with the addition of the new editor, to inform them of the change. Another \texttt{BRD\_USERS} message is sent to all users of the previous board, with the omission of the removed editor, to inform them of the change.
\item A sequence of \texttt{STROKE} messages are sent to the client to recreate the current state of the selected whiteboard.
\end{enumerate}

\paragraph{\texttt{BRD\_REQ}} When the client wants to create a new board, it send a \texttt{BRD\_REQ} message with a properly formatted \texttt{NAME}.
\begin{enumerate}
\item The server initializes a new \texttt{MasterBoard} object with relevant properties and adds it to its list of boards.
\item A \texttt{BRD\_INFO} message for the new board is sent to all connected clients to inform them of the newly available whiteboard. The clients add this board to their list of available boards.
\end{enumerate}

\paragraph{\texttt{BRD\_ALL}} This is often called by the client upon initialization. It prompts a series of \texttt{BRD\_INFO} replies for all existing boards.

\paragraph{\texttt{BRD\_DEL}} When the client wants to remove a board, it send a \texttt{BRD\_DEL} message to the server.
\begin{enumerate}
\item If the provided \texttt{BOARD\_ID} exists, the server removes the designated board from its central list of boards.
\item The \texttt{BRD\_DEL} message is forwarded to all clients to signify that the board is no longer available. The clients remove this board from their list of available boards.
\end{enumerate}

\paragraph{\texttt{BRD\_CLR}} When the client wants to clear a board, it sends a \texttt{BRD\_CLR} message to the server.
\begin{enumerate}
\item If the provided \texttt{BOARD\_ID} exists, the server clears the contents of the board, all associated strokes.
\item The \texttt{BRD\_CLR} message is forwarded to all clients to signify that the board has been cleared.
\end{enumerate}

\paragraph{\texttt{USER\_REQ}} Upon entering a username in the client application, a \texttt{USER\_REQ} message with a properly formatted \texttt{USER\_NAME} will be sent to the server to request the desired username.
\begin{enumerate}
\item The requested username is checked against existing username. If the name already exists, a name is generated using the new board's ID number.
\item A \texttt{User} object with relevant properties is created to represent the new client.
\item A \texttt{USER\_INIT} message is sent to the client to inform it of its acquired username.
\end{enumerate}

Note that \texttt{BRD\_USERS}, \texttt{BRD\_INFO}, and \texttt{USER\_INIT} are used solely as replies and are included in the descriptions above.


\subsection{Data Transport}
For a detailed explanation of the communication method between the clients and the server, please see the "Threads and Queues" section.